Real Estate

### 2022-11-18 src/demo_data_wrangling.Rmd
updating score for MSAs without occupancy rates

MSAs <-metro_house_prices %>% 
  filter(popestimate2021 >= 500000)
```

@@

gs4_create("updated_MSAs")
sheet_write(MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSAs")
```

@@

```{r MSAs seperate into cities}
#  split up MSAs into cities to collect utility, rent, and landlord data.
msa_seperated <- metro_house_prices %>% 
  separate_rows(MSA, sep = "-") %>% 
  rename(city = MSA) %>% # seperate MSA cities into their own columns
 drop_na(city)

gs4_create("updated_MSAs")
sheet_write(msa_seperated, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "cities")

>>>>

MSAs <-metro_house_prices %>% 
  filter(popestimate2021 >= 500000) %>% 
  distinct(cbsa, .keep_all = TRUE) # remove duplicate portland row
```

@@

#gs4_create("updated_MSAs")
sheet_write(MSAs, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "MSAs")
```

@@

```{r MSAs seperate into cities}
#  split up MSAs into cities to collect utility, rent, and landlord data.
msa_seperated <- MSAs %>% 
  separate_rows(MSA, sep = "-") %>% 
  rename(city = MSA) %>% # separate MSA cities into their own columns
  filter(! city == "")

#gs4_create("updated_MSAs")
sheet_write(msa_seperated, ss = "https://docs.google.com/spreadsheets/d/1lmTpSDwVANxdAg5sW87Q8l_gtig7M0EjPPJ6J5_7dwg/edit#gid=0", sheet = "cities")

### 2022-11-18 src/real_estate.Rmd
saved data to google sheets in r folder on google drive
added function to print datatable. Double checked landlord policy
updating score for MSAs without occupancy rates

library(tidyverse)
library(janitor)
library(googlesheets4)

@@

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(MSA_realEstate_data) %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(avg_rent_to_income = (avg_rent_2021*12)/median_income) %>% 
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

@@

```{r}
MSA_n_data_avg_rent <- n_MSA_realEstate %>% 
  select(cbsa, MSA, n_pop_change, n_employ_change, n_avg_occupancy, n_avg_rent_change, n_avg_rent_to_income_inverted, n_income_to_homeprice)

MSA_n_data_med_rent <- n_MSA_realEstate %>% 
  select(cbsa, MSA, n_pop_change, n_employ_change, n_avg_occupancy, n_med_rent_change, n_med_rent_to_income_inverted, n_income_to_homeprice)
```

@@

#make real_estate_score column
n_MSA_realEstate <- n_MSA_realEstate %>% 
  select(cbsa, MSA, starts_with("n_")) %>% 
  mutate(real_estate_score = MSA) %>% #this is just a place holder
  mutate(RE_score_avg_rent = MSA)
```

@@

>>>>

library(tidyverse)
library(janitor)
library(googlesheets4)
library(DT)

@@

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(MSA_realEstate_data) %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(avg_rent_to_income = (avg_rent_2021*12)/median_income) %>% 
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

@@

@@

#make real_estate_score column
n_MSA_realEstate <- n_MSA_realEstate %>% 
  select(cbsa, MSA, starts_with("n_"), ) %>% 
  mutate(real_estate_score = MSA) %>% #this is just a place holder
  mutate(RE_score_avg_rent = MSA)
```

@@

```{r}
# create downloadable data table
datatable(n_MSA_realEstate, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)
```

```{r}
#write result to google sheet
# gs4_create("MSA_real_estate_score")
sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "n_MSA_realEstate")
```
